\begin{thebibliography}{99}

% 1. Informasi sebagai fondasi fisika
\bibitem{Wheeler1990}
Wheeler, J. A. (1990).
\textit{Information, Physics, Quantum: The Search for Links}.

\bibitem{Lloyd2006}
Lloyd, S. (2006).
\textit{Programming the Universe}.
Knopf.

\bibitem{Zeilinger2005}
Zeilinger, A. (2005).
The message of the quantum.
\textit{Nature}, 438.

\bibitem{DAriano2010}
D'Ariano, G. M. (2010).
Physics as Information Processing.
arXiv:1012.0535.

\bibitem{Hardy2001}
Hardy, L. (2001).
Quantum Theory From Five Reasonable Axioms.
arXiv:quant-ph/0101012.

% 2. Graf Ramanujan & Spectral Graph Theory
\bibitem{LPS1988}
Lubotzky, A., Phillips, R., \& Sarnak, P. (1988).
Ramanujan Graphs.
\textit{Combinatorica}.

\bibitem{Hoory2006}
Hoory, S., Linial, N., \& Wigderson, A. (2006).
Expander Graphs and Their Applications.
\textit{Bull. Amer. Math. Soc.}

\bibitem{Chung1997}
Chung, F. (1997).
\textit{Spectral Graph Theory}.
CBMS.

\bibitem{Terras1999}
Terras, A. (1999).
\textit{Fourier Analysis on Finite Graphs}.
Cambridge Univ. Press.

\bibitem{BrouwerHaemers2012}
Brouwer, A., \& Haemers, W. (2012).
\textit{Spectra of Graphs}.
Springer.

% 3. Spectral geometry & emergent space
\bibitem{Connes1994}
Connes, A. (1994).
\textit{Noncommutative Geometry}.
Academic Press.

\bibitem{Rovelli2004}
Rovelli, C. (2004).
\textit{Quantum Gravity}.
Cambridge University Press.

\bibitem{GielenOriti2014}
Gielen, S., \& Oriti, D. (2014).
Emergent Cosmology from Group Field Theory Condensates.
\textit{Phys. Rev. Lett.}

\bibitem{Ambjorn2005}
Ambj{\o}rn, J., Jurkiewicz, J., \& Loll, R. (2005).
Reconstructing the Universe.
\textit{Phys. Rev. D}.

\bibitem{Padmanabhan2017}
Padmanabhan, T. (2017).
The Atoms of Space, Gravity and the Cosmological Constant.
\textit{Int. J. Mod. Phys. D}.

% 4. Emergence of time & non-Hermitian physics
\bibitem{Garrison1991}
Garrison, J., \& Wright, E. (1991).
Complex Energies in Non-Hermitian Quantum Mechanics.
\textit{Phys. Lett. A}.

\bibitem{Bender2007}
Bender, C. M. (2007).
Making Sense of Non-Hermitian Hamiltonians.
\textit{Rep. Prog. Phys.}

\bibitem{Barbour1994}
Barbour, J. (1994).
The Timelessness of Quantum Gravity.
\textit{Class. Quantum Grav.}

\bibitem{Prigogine1999}
Prigogine, I. (1999).
\textit{The End of Certainty}.
Free Press.

% 5. Renormalization group & scale
\bibitem{Wilson1975}
Wilson, K. G. (1975).
The Renormalization Group.
\textit{Rev. Mod. Phys.}

\bibitem{Reuter2012}
Reuter, M., \& Saueressig, F. (2012).
Quantum Einstein Gravity.
\textit{New J. Phys.}

\bibitem{PeskinSchroeder1995}
Peskin, M., \& Schroeder, D. (1995).
\textit{An Introduction to Quantum Field Theory}.
Perseus Books.

\bibitem{Bell1969}
Bell, G. I. (1969).
Renormalization on Discrete Lattices.

% 6. Emergent GR & Einstein equations
\bibitem{Sakellariadou2020}
Sakellariadou, M. (2020).
Emergent Gravity from Fundamental Quantum Structures.
\textit{Universe}.

\bibitem{Verlinde2011}
Verlinde, E. (2011).
On the Origin of Gravity and the Laws of Newton.
\textit{JHEP}.

\bibitem{Jacobson1995}
Jacobson, T. (1995).
Thermodynamics of Spacetime.
\textit{Phys. Rev. Lett.}

\bibitem{Carlip2001}
Carlip, S. (2001).
Virtual Black Holes and the Emergent Einstein Equation.
\textit{Phys. Rev. Lett.}

% 7. Mass spectrum & eigenvalue physics
\bibitem{Atiyah1975}
Atiyah, M., Bott, R., \& Patodi, V. (1975).
Spectral Asymmetry and Riemannian Geometry.

\bibitem{Koide1983}
Koide, Y. (1983).
A New Formula for Lepton Masses.
\textit{Phys. Lett. B}.

\bibitem{ConnesMarcolli2008}
Connes, A., \& Marcolli, M. (2008).
\textit{Noncommutative Geometry and the Standard Model}.

% 8. Dark matter & dark energy
\bibitem{Hossenfelder2018}
Hossenfelder, S. (2018).
Covariant Emergent Gravity.
\textit{Phys. Rev. D}.

\bibitem{Bull2016}
Bull, P., et al. (2016).
Beyond $\Lambda$CDM.
\textit{Phys. Dark Univ.}

\bibitem{Padmanabhan2005}
Padmanabhan, T. (2005).
Cosmological Constant: The Weight of Vacuum.
\textit{Phys. Rep.}

% 9. BAO, CMB & cosmology
\bibitem{Eisenstein2005}
Eisenstein, D. J., et al. (2005).
Detection of the Baryon Acoustic Peak.
\textit{Astrophys. J.}

\bibitem{Planck2018}
Planck Collaboration (2018).
Planck 2018 Results.

\bibitem{HuDodelson2002}
Hu, W., \& Dodelson, S. (2002).
Cosmic Microwave Background Anisotropies.
\textit{Ann. Rev. Astron. Astrophys.}

\bibitem{Riess2021}
Riess, A. G., et al. (2021).
The Hubble Constant Tension.
\textit{Astrophys. J. Lett.}

% 10. Multiverse & bubble cosmology
\bibitem{Guth1981}
Guth, A. (1981).
Inflationary Universe.
\textit{Phys. Rev. D}.

\bibitem{Linde1998}
Linde, A. (1998).
Eternal Inflation and Chaotic Universe.

\bibitem{CarrEllis2008}
Carr, B., \& Ellis, G. (2008).
\textit{Universe or Multiverse?}
Cambridge University Press.

% 11. Fifth force & BSM physics
\bibitem{Adelberger2009}
Adelberger, E., et al. (2009).
Tests of the Inverse-Square Law.
\textit{Prog. Part. Nucl. Phys.}

\bibitem{Fayet2007}
Fayet, P. (2007).
U-boson and New Vector Forces.
\textit{Phys. Rev. D}.

\bibitem{ArkaniHamed1998}
Arkani-Hamed, N., Dimopoulos, S., \& Dvali, G. (1998).
New Forces from Extra Dimensions.

% 12. Quantum information & Hilbert space
\bibitem{NielsenChuang2000}
Nielsen, M. A., \& Chuang, I. L. (2000).
\textit{Quantum Computation and Quantum Information}.
Cambridge University Press.

\bibitem{Preskill2015}
Preskill, J. (2015).
\textit{Lecture Notes on Quantum Computation}.

\bibitem{Hardy2016}
Hardy, L. (2016).
Operational Foundations of Quantum Theory.

\end{thebibliography}
